### 2020-10-28 doc/manuscript/Dow_manuscript.Rmd
add HF description

Measurements were timed to begin before beginning of spring growth and to continue through the cessation of growth in the fall. *(Give specifics, which may vary by site. Note that we adjusted timing at SCBI when we observed early leaf-out of understory shrubs/herbs)* The median start date was 4/14, which was adjusted when early leaf-out of understory shrubs and herbs was observed, and the median end date was 10/17, coinciding with late senesence in the plot. The earliest start date for intraannual measurements was in 2020 on March 30, and the latest end date was on 2012 on November 26.

The dendroband data were manually cleaned by visual inspection. Three classes of mistakes: 1. Error in measurement - weekly measurement was drastically different from previous week and following week. Mistakes were removed from the raw data before modeling. (## cases) 2. Band slip or stuck - measurements freeze and remain unchanged until sudden jump followed by normal growth pattern. Band slips were sometimes followed by measurements indicating the band became stuck; in these cases, the tree was thrown out for the corresponding year. In cases where the slip was followed by normal growth, the initial slip-point was removed, and the following points were shifted down to the pre-slip level. (## cases) 3. Other - cases where data was clearly wrong but with unknown causes. This seemed to happen mostly in 2011, when the program was just beginning. In cases where several measurements were an issue, they were simply removed. If there were no clear solutions to ‘fix’ the data, the entire year was removed from the analysis.

Daily maximum temperatures (TMAX) were obtained from a meteorological tower adjacent to the SCBI ForestGEO plot (data can be found on the ForestGEO climate portal, DOI on github). The R package climpact was used to plot TMAX for visual inspection (# points removed?). Climpact also identifies readings that were greater than 3 standard deviations away from yearly means. These points were labeled as outliers and removed from the dataset. Gaps in the SCBI meteorological tower data were subsequently filled using TMAX readings obtained from a National Center for Environmental Information (NCEI) weather station located in Front Royal, Virginia (https://www.ncdc.noaa.gov/cdo-web/datasets/GHCND/stations/GHCND:USC00443229/detail).

>>>>

Measurements were timed to begin before beginning of spring growth and to continue through the cessation of growth in the fall. *(Give specifics, which may vary by site. Note that we adjusted timing at SCBI when we observed early leaf-out of understory shrubs/herbs)* At SCBI, the median start date was 4/14, which was adjusted when early leaf-out of understory shrubs and herbs was observed, and the median end date was 10/17, coinciding with late senescence in the plot. The earliest start date for intraannual measurements at SCBI was March 30, 2020 and the latest end date was November 26, 2012. Harvard Forest measurements took place at a similar time, with the median start date of 4/23 and median end date of 10/30. 1998 was an anomalous year where initial measurements were taken on 1/5, but not taken again until 4/15. The latest end date was November 11, 2002.

The dendroband data were manually cleaned by visual inspection. Three classes of mistakes: 1. Error in measurement - weekly measurement was drastically different from previous week and following week. Mistakes were removed from the raw data before modeling. 2. Band slip or stuck - measurements freeze and remain unchanged until sudden jump followed by normal growth pattern. Band slips were sometimes followed by measurements indicating the band became stuck; in these cases, the tree was thrown out for the corresponding year. In cases where the slip was followed by normal growth, the initial slip-point was removed, and the following points were shifted down to the pre-slip level. 3. Other - cases where data was clearly wrong but with unknown causes. In cases where several measurements were an issue, they were simply removed. If there were no clear solutions to ‘fix’ the data, the entire year was removed from the analysis.

Daily maximum temperatures (TMAX) for SCBI were obtained from a meteorological tower adjacent to the SCBI ForestGEO plot (data can be found on the ForestGEO climate portal, DOI on github). Daily maximum temperatures at Harvard Forest were obtained from the Harvard Forest weather station (Boose & Gould, 2004; Boose, 2018).  The R package climpact () was used to plot TMAX for visual inspection. Climpact also identifies readings that were greater than 3 standard deviations away from yearly means. These points were labeled as outliers and removed from the dataset. Gaps in the SCBI meteorological tower data were subsequently filled using TMAX readings obtained from a National Center for Environmental Information (NCEI) weather station located in Front Royal, Virginia (https://www.ncdc.noaa.gov/cdo-web/datasets/GHCND/stations/GHCND:USC00443229/detail).
